Astr 513: Final Project

### 2016-05-05 Final_Project/ms.tex
Added some whitespace to make equations easier to read.

\subsection{The JLA sample}
\label{sec:betoule}
\citet{betoule2014} define their ``joint light-curve analysis'' (JLA) 
sample, a total of 740 \snea{} compiled from a combination of 
data from the Sloan Digital Sky 
Survey (SDSS), the Supernova Legacy Survey (SNLS), the Hubble Space 
Telescope (HST), and several other experiments. The 613 SDSS \snea{} 
come from the SDSS supernova survey (SDSS-SNS) of SDSS-II \citep{sako2014} 
with redshifts of $0.05 < z < 0.4$, and 374 of them are selected from those 
that were confirmed in the spectroscopic follow-up. The rest of the 
\sne{} come from the compilation by \citet{conley2011}, which includes 
242 spectroscopically confirmed \snea{} from SNLS at redshifts $0.2 < z < 1$, 
14 \snea{} from HST at very high redshifts of $0.7 < z < 1.4$, and 123 
low redshift ($z < 0.08$) from a variety of sources. We made use of the 
publicly available distance moduli, redshifts, and covariance 
matrix provided by \citet{betoule2014}\footnote{Data is available 
for download via 
\url{http://cdsarc.u-strasbg.fr/viz-bin/qcat?J/A+A/568/A22}}. 
We note that the distance moduli are binned, so that they are only computed and 
given on a grid of redshifts. The distance modulus can then be well 
approximated by a piecewise linear function of $\log(z)$ that is given 
in a single redshift bin $z_b \le z < z_{b+1}$ as 
\citep[see][Appendix E.1]{betoule2014}
\begin{equation}
\label{eq:binnedMu}
\bar{\mu}(z) = (1 - \alpha) \mu_b + \alpha \mu_{b+1},
\end{equation}
where they have defined 
$\alpha = \log\left(\frac{z}{z_b}\right)/\log\left(\frac{z_{b+1}}{z_b}\right)$ 
and $\mu_i$ is the distance modulus at control point $z_i$. 
Therefore, the data and covariance matrix are merely given at the 
31 control points log-spaced in the redshift range $0.01 < z < 1.3$.

@@

\subsection{Covariances}
\label{sec:covariances}
The publicly released covariance matrix from \citet{rest2014} 
is comprised of systematic and statistical errors for the final sample. 
The total covariance matrix is defined as 
$\mathbf{C}= \mathbf{D}_{\textnormal{stats}}+\mathbf{C}_{\textnormal{sys}}$. The 
statistical matrix has only diagonal elements, which take into account errors 
in light curve fit parameters and intrinsic scatter due to luminosity and 
color variation. These statistical errors are added in quadratures. The 
systematics, defined in \citet{scolnic2014}, include flux calibration errors, 
selection effects, light curve fitting, host galaxy relations, Milky Way 
extinction corrections, and coherent velocity flows. Calculating the systematic 
matrix involves varying a systematic parameter and determining the difference 
between the original distance and the new distance. The systematic covariance 
between {\sn} $i$ and {\sn} $j$ is 
\begin{equation}
\label{eq:RestCovMatrix}
C_{ij,\textnormal{sys}}=\displaystyle \sum_{k = 1}^K \left(\frac{\partial \mu_i}{\partial S_k}\right)\left(\frac{\partial \mu_j}{\partial S_k}\right)\left(\sigma_{S_k}\right)^2,
\end{equation}
where the summation is over $K$ systematics, $\sigma_{S_k}$ is each 
systematic error magnitude, and $\partial \mu$ is the difference in 
distance modulus from varying a systematic parameter.

\citet{betoule2014} choose to first calculate the covariance 
matrix $\mathbf{C}_{\mathbf{\eta}}$ of the 3 light curve fit-parameters $\mathbf{\eta}$ 
by adding the individual covariance matrices they get from statistical errors propagated 
from light-curve fit uncertainties and 7 sources of systematic uncertainties 
they consider from their photometric calibartion, light-curve model uncertainty, 
bias correction, a mass step uncertainty introduced when correcting magnitudes due 
to host mass, Galactic extinction, peculiar velocities, and contamination by 
non-\sna{}. They then propagate these errors to the distance modulus and use three 
terms to account for uncertainty in cosmological redshift by the equation
\begin{equation}
\label{eq:BetouleCovMatrix}
\mathbf{C} = \mathbf{A} \mathbf{C}_\mathbf{\eta} \mathbf{A}^\dag + \mathrm{diag}\left(\frac{5 \sigma_z}{z \log 10}\right)^2 + \mathrm{diag}\left(\sigma_{\mathrm{lens}}^2\right) + \mathrm{diag}\left(\sigma_{\mathrm{coh}}^2\right),
\end{equation}
where $\mathbf{A}$ is the matrix of coefficients in the distance modulus, 
and redshift uncertainty is accounted for in the first diagonal term due to 
peculiar velocities, the second due to magnitude variation from gravitational lensing, 
and the third from intrinsic \sn{} magnitude variation not included in other terms. 
They use $c \sigma_z = 150 \mathrm{km}\, \mathrm{s}^{-1}$, 
$\sigma_{\mathrm{lens}} = 0.055 \times z$, and a value of $\sigma_{\mathrm{coh}}$ for each 
survey found using a restricted log-likelihood technique. The elements of the covariance 
matrix available online are calculated on the redshift control points $z_b$ as 
described in \S \ref{sec:betoule}.

@@

The correlation matrix is calculated from the covariance matrix according to
\begin{equation}
  \label{eq:correlation}
  Corr_{i,j} = \frac{\mathbf{C}_{i,j}}{\sqrt{\mathbf{C}_{i,i}}\sqrt{\mathbf{C}_{j,j}}}.
\end{equation}
Note that the correlation of a variable with itself is 1 by definition.
The correlation matrices for both of our data sets can be seen below in
figures \autoref{fig:cmatrix_betoule} and \autoref{fig:cmatrix_rest}.

@@

\section{Distance Modulus Model}
\label{sec:model}
The main objective of \citet{betoule2014} and \citet{rest2014} is 
to place constraints on cosmological parameters $\Omega_M$, $\Omega_\Lambda$, 
$\Omega_k$, and $w$. In order to so, we must examine the relationship between distance 
modulus and cosmology. The distance modulus $\mu$ is determined from 
the luminosity distance $d_L$ by 
\begin{equation}
\label{eq:mu_dL}
\mu=5 \log\left(\frac{d_L}{1 \mathrm{Mpc}}\right)+25. 
\end{equation}
The luminosity distance depends on the cosmology by 
\begin{equation}
\label{eq:luminosityDistance}
d_L(z; w, \Omega_M, \Omega_\Lambda, H_0)= (1+z)|\Omega_k|^{-1/2}\mathcal{S}_{k}\displaystyle\left[c|\Omega_k|^{1/2}\int_{0}^{z}\frac{dz'}{H(z')}\right],
\end{equation}
where $\Omega_k \equiv 1-\Omega_M-\Omega_{\Lambda}$. The function 
$\mathcal{S}_{k}$ depends on the curvature density $\Omega_k$ in that 
\[
    \mathcal{S}_{k}= 
\begin{cases}
    \text{sin}(x),&  \Omega_k < 0\\
    x,& \Omega_k = 0\\
    \text{sinh}(x),& \Omega_k > 0
\end{cases}.
\]
The function $H(z)$ is expressed as 
\begin{equation}
\label{eq:hubbleFunction}
H(z)= H_0\displaystyle\left[\Omega_M(1+z)^3+\Omega_{\Lambda}(1+z)^{3(1+w)}+\Omega_k(1+z)^2\right]^{1/2},
\end{equation} where the Hubble constant $H_0$ can be written in terms of a 
dimensionless Hubble parameter $h$ by $H_0=100h \mathrm{km\ s^{-1}Mpc^{-1}}$.

@@

\subsection{Likelihood}
\label{sec:likelihood}
Using the above model for the distance modulus allows us to write a
probability that a model can generate our data:
%
\begin{eqnarray}
  \label{eq:likelihood}
  P(\mu(z)|\mathrm{cosmo}) = \frac{1}{\sqrt{2\pi|\mathbf{C}|}}\exp
  \left[-\frac{1}{2}& \\ \sum_{i,j=0}^{N_{samples}}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_i 
    \mathbf{C}^{-1}_{i,j}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_j\right]
\end{eqnarray}
%
where $\mathbf{C}$ is the full covariance matrix between all of the $N_{samples}$ 
in the data, and $\mathrm{cosmo}$ represents the cosmology that we are investigating
(e.g. flat $\Lambda$CDM, $wCDM$, etc.). 
This is not the full picture though, since we must employ
Bayes' Theorum in order to include any priors on our choice of 
parameters and recover the posterior distribution function of our
parameters given our data
%
\begin{equation}
  \label{eq:posterior}
  P(\mathrm{cosmo}|\mu(z)) = P_{prior}(\mathrm{cosmo}) P(\mu(z)|\mathrm{cosmo}).
\end{equation}
%
For this work we investigate a combination of priors. \Cosmosis is implemented
in a way that requires the user to provide bounds on the parameter space, which we
detail for each parameter in Table~XXX
\TM{Make this table. Just look at the cosmology.ini files.}
Within those bounds, we investigated using a flat prior $P_{prior}=1$ and
\Planck priors.

>>>>

\subsection{The JLA sample}
\label{sec:betoule}
\citet{betoule2014} define their ``joint light-curve analysis'' (JLA) 
sample, a total of 740 \snea{} compiled from a combination of 
data from the Sloan Digital Sky 
Survey (SDSS), the Supernova Legacy Survey (SNLS), the Hubble Space 
Telescope (HST), and several other experiments. The 613 SDSS \snea{} 
come from the SDSS supernova survey (SDSS-SNS) of SDSS-II \citep{sako2014} 
with redshifts of $0.05 < z < 0.4$, and 374 of them are selected from those 
that were confirmed in the spectroscopic follow-up. The rest of the 
\sne{} come from the compilation by \citet{conley2011}, which includes 
242 spectroscopically confirmed \snea{} from SNLS at redshifts $0.2 < z < 1$, 
14 \snea{} from HST at very high redshifts of $0.7 < z < 1.4$, and 123 
low redshift ($z < 0.08$) from a variety of sources. We made use of the 
publicly available distance moduli, redshifts, and covariance 
matrix provided by \citet{betoule2014}\footnote{Data is available 
for download via 
\url{http://cdsarc.u-strasbg.fr/viz-bin/qcat?J/A+A/568/A22}}. 
We note that the distance moduli are binned, so that they are only computed and 
given on a grid of redshifts. The distance modulus can then be well 
approximated by a piecewise linear function of $\log(z)$ that is given 
in a single redshift bin $z_b \le z < z_{b+1}$ as 
\citep[see][Appendix E.1]{betoule2014}
%
\begin{equation}
\label{eq:binnedMu}
\bar{\mu}(z) = (1 - \alpha) \mu_b + \alpha \mu_{b+1},
\end{equation}
%
where they have defined 
$\alpha = \log\left(\frac{z}{z_b}\right)/\log\left(\frac{z_{b+1}}{z_b}\right)$ 
and $\mu_i$ is the distance modulus at control point $z_i$. 
Therefore, the data and covariance matrix are merely given at the 
31 control points log-spaced in the redshift range $0.01 < z < 1.3$.

@@

\subsection{Covariances}
\label{sec:covariances}
The publicly released covariance matrix from \citet{rest2014} 
is comprised of systematic and statistical errors for the final sample. 
The total covariance matrix is defined as 
$\mathbf{C}= \mathbf{D}_{\textnormal{stats}}+\mathbf{C}_{\textnormal{sys}}$. The 
statistical matrix has only diagonal elements, which take into account errors 
in light curve fit parameters and intrinsic scatter due to luminosity and 
color variation. These statistical errors are added in quadratures. The 
systematics, defined in \citet{scolnic2014}, include flux calibration errors, 
selection effects, light curve fitting, host galaxy relations, Milky Way 
extinction corrections, and coherent velocity flows. Calculating the systematic 
matrix involves varying a systematic parameter and determining the difference 
between the original distance and the new distance. The systematic covariance 
between {\sn} $i$ and {\sn} $j$ is 
%
\begin{equation}
\label{eq:RestCovMatrix}
C_{ij,\textnormal{sys}}=\displaystyle \sum_{k = 1}^K \left(\frac{\partial \mu_i}{\partial S_k}\right)\left(\frac{\partial \mu_j}{\partial S_k}\right)\left(\sigma_{S_k}\right)^2,
\end{equation}
%
where the summation is over $K$ systematics, $\sigma_{S_k}$ is each 
systematic error magnitude, and $\partial \mu$ is the difference in 
distance modulus from varying a systematic parameter.

\citet{betoule2014} choose to first calculate the covariance 
matrix $\mathbf{C}_{\mathbf{\eta}}$ of the 3 light curve fit-parameters $\mathbf{\eta}$ 
by adding the individual covariance matrices they get from statistical errors propagated 
from light-curve fit uncertainties and 7 sources of systematic uncertainties 
they consider from their photometric calibartion, light-curve model uncertainty, 
bias correction, a mass step uncertainty introduced when correcting magnitudes due 
to host mass, Galactic extinction, peculiar velocities, and contamination by 
non-\sna{}. They then propagate these errors to the distance modulus and use three 
terms to account for uncertainty in cosmological redshift by the equation
%
\begin{equation}
\label{eq:BetouleCovMatrix}
\mathbf{C} = \mathbf{A} \mathbf{C}_\mathbf{\eta} \mathbf{A}^\dag + \mathrm{diag}\left(\frac{5 \sigma_z}{z \log 10}\right)^2 + \mathrm{diag}\left(\sigma_{\mathrm{lens}}^2\right) + \mathrm{diag}\left(\sigma_{\mathrm{coh}}^2\right),
\end{equation}
%
where $\mathbf{A}$ is the matrix of coefficients in the distance modulus, 
and redshift uncertainty is accounted for in the first diagonal term due to 
peculiar velocities, the second due to magnitude variation from gravitational lensing, 
and the third from intrinsic \sn{} magnitude variation not included in other terms. 
They use $c \sigma_z = 150 \mathrm{km}\, \mathrm{s}^{-1}$, 
$\sigma_{\mathrm{lens}} = 0.055 \times z$, and a value of $\sigma_{\mathrm{coh}}$ for each 
survey found using a restricted log-likelihood technique. The elements of the covariance 
matrix available online are calculated on the redshift control points $z_b$ as 
described in \S \ref{sec:betoule}.

@@

The correlation matrix is calculated from the covariance matrix according to
%
\begin{equation}
  \label{eq:correlation}
  Corr_{i,j} = \frac{\mathbf{C}_{i,j}}{\sqrt{\mathbf{C}_{i,i}}\sqrt{\mathbf{C}_{j,j}}}.
\end{equation}
%
Note that the correlation of a variable with itself is 1 by definition.
The correlation matrices for both of our data sets can be seen below in
figures \autoref{fig:cmatrix_betoule} and \autoref{fig:cmatrix_rest}.

@@

\section{Distance Modulus Model}
\label{sec:model}
The main objective of \citet{betoule2014} and \citet{rest2014} is 
to place constraints on cosmological parameters $\Omega_M$, $\Omega_\Lambda$, 
$\Omega_k$, and $w$. In order to so, we must examine the relationship between distance 
modulus and cosmology. The distance modulus $\mu$ is determined from 
the luminosity distance $d_L$ by 
%
\begin{equation}
  \label{eq:mu_dL}
  \mu=5 \log\left(\frac{d_L}{1 \mathrm{Mpc}}\right)+25. 
\end{equation}
%
The luminosity distance depends on the cosmology by 
%
\begin{equation}
\label{eq:luminosityDistance}
d_L(z; w, \Omega_M, \Omega_\Lambda, H_0)= (1+z)|\Omega_k|^{-1/2}\mathcal{S}_{k}\displaystyle\left[c|\Omega_k|^{1/2}\int_{0}^{z}\frac{dz'}{H(z')}\right],
\end{equation}
%
where $\Omega_k \equiv 1-\Omega_M-\Omega_{\Lambda}$. The function 
$\mathcal{S}_{k}$ depends on the curvature density $\Omega_k$ in that 
\[
    \mathcal{S}_{k}= 
\begin{cases}
    \text{sin}(x),&  \Omega_k < 0\\
    x,& \Omega_k = 0\\
    \text{sinh}(x),& \Omega_k > 0
\end{cases}.
\]
The function $H(z)$ is expressed as 
%
\begin{equation}
  \label{eq:hubbleFunction}
  H(z)= H_0\displaystyle\left[\Omega_M(1+z)^3+\Omega_{\Lambda}(1+z)^{3(1+w)}+\Omega_k(1+z)^2\right]^{1/2},
\end{equation}
%
where the Hubble constant $H_0$ can be written in terms of a 
dimensionless Hubble parameter $h$ by $H_0=100h \mathrm{km\ s^{-1}Mpc^{-1}}$.

@@

\subsection{Likelihood}
\label{sec:likelihood}
Using the above model for the distance modulus allows us to write a
probability that a model can generate our data:
%
\begin{eqnarray}
  \label{eq:likelihood}
  P(\mu(z)|\mathrm{cosmo}) = \frac{1}{\sqrt{2\pi|\mathbf{C}|}}\exp
  \left[-\frac{1}{2} \sum_{i,j=0}^{N_{samples}}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_i 
    \mathbf{C}^{-1}_{i,j}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_j\right]
\end{eqnarray}
%
where $\mathbf{C}$ is the full covariance matrix between all of the $N_{samples}$ 
in the data, and $\mathrm{cosmo}$ represents the cosmology that we are investigating
(e.g. flat $\Lambda$CDM, $wCDM$, etc.). 
This is not the full picture though, since we must employ
Bayes' Theorum in order to include any priors on our choice of 
parameters and recover the posterior distribution function of our
parameters given our data
%
\begin{equation}
  \label{eq:posterior}
  P(\mathrm{cosmo}|\mu(z)) = P_{prior}(\mathrm{cosmo}) P(\mu(z)|\mathrm{cosmo}).
\end{equation}
%
For this work we investigate a combination of priors. \Cosmosis is implemented
in a way that requires the user to provide bounds on the parameter space, which we
detail for each parameter in \autoref{tab:priors}
Within those bounds, we investigated using a flat prior $P_{prior}=1$ and
\Planck priors.

\begin{table}
  \centering
  \caption{Bounds on cosmological parameters 
    entering \autoref{eq:posterior}. Our sampler requires
    bounds in order to both populate the parameter space with
    parallelized samplers and to calculate an appropriate width
    for the step size.}
  \label{tab:priors}
  \begin{tabular}{lll}
    Parameter & Description & Bounds \\ \hline
    $h$ & Hubble constant$/100$ & $[0,2]$ \\
    $\Omega_M$ & Matter fraction at $z=0$ & $[0,1]$ \\
    $w$ & Dark energy E.O.S. & $[-2,0]$\\
  \end{tabular}
\end{table}
